Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-06-22 paper.tex
Final edits

%Review style settings
%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

@@

For example, iChemLabs released the ChemDoodle Web Components library under the GPL v3
license, based on the upcoming HTML5 Open Standard. It allows making web and mobile
interfaces for chemical content. The project is already being adopted by others,
including iBabel~\cite{iBabel}, ChemSpotlight~\cite{chemspotlight} and the RSC ChemSpider~\cite{chemspider_chemdoodle}.

@@

Despite this positive news, little chemical data has become
available from the traditional chemical fields of organic,
inorganic, solid state chemistry. Table~2 lists a few notable
exceptions: Metamolecular's Chempedia~\cite{Chempedia} (a discontinued substance registration
and naming service whose raw data can still be accessed),
CrystalEye~\cite{WebCrystalEye},
and the Open Notebook Science Solubility
data~\cite{ONS2010}. There is also data available using licenses
not compatible with the Panton Principles, but where the user
is allowed to modify and redistribute the data. A new data
set in this category is the data from the ChEMBL database,
which is available under the Creative Commons Share-Alike
Attribution license~\cite{Overington2009}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
   The inital layout of the manuscript grew from discussions between
NMOB, RG and ELW. All of the authors contributed sections to this
article, and approved the final manuscript.

>>>>

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

@@

For example, iChemLabs released the ChemDoodle Web Components library under the GPL v3
license, based on the upcoming HTML5 Open Standard. It allows making web and mobile
interfaces for chemical content. The project is already being adopted by others,
including iBabel~\cite{iBabel}, ChemSpotlight~\cite{chemspotlight} and the RSC ChemSpider~\cite{chemspider,chemspider_chemdoodle}.

@@

Despite this positive news, little chemical data has become
available from the traditional chemical fields of organic,
inorganic, solid state chemistry. Table~2 lists a few notable
exceptions: Metamolecular's Chempedia~\cite{Chempedia} (a discontinued substance registration
and naming service whose raw data can still be accessed),
CrystalEye~\cite{WebCrystalEye},
and the Open Notebook Science Solubility
data~\cite{ONS2010}. There is also data available using licenses
not compatible with the Panton Principles, but where the user
is allowed to modify and redistribute the data. A new data
set in this category is the data from the ChEMBL database,
which is available under the Creative Commons Share-Alike
Attribution license~\cite{Overington2009}. The RSC ChemSpider database~\cite{chemspider}, although
not fully Open, also hosts Open Data; for example, spectral data when deposited
can be marked as Open.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors' contributions}
The overall layout of the manuscript grew from discussions between NMOB, RG and ELW. The authorship of the paper is drawn from those people connected with fully Open Data/Standards/Source (OSI-compliant or OKF-compliant) projects associated with the Blue Obelisk. There are a large number of people contributing to these projects and because those projects are published in their own right it is not appropriate to include all their developers by default. We invited a number of `project gurus' who have been active in promoting the Blue Obelisk, to be authors on this paper and most have accepted and contributed.
